---
title: "HCQO validation and calculation codes v3.0"
author: "Aksel Betkas, Kadri-Ann Kallas, Elise Barattini"
date: "2023/24"
output: html_document
---

# This is a Rmarkdown code to be run on data files from the OECD HCQO data collection. 

### Here we set up R and if necessary, download the R libraries you need to run the code
To run the codes, you will need to 
  1. have downloaded the folder called HCQO_validation
  2. install the necessary R packages (R should let you know which ones are necessary)
  3. copy .csv files you wish to analyse into the folder HCQO_validation\data*
  4. Run the code in the gray boxes without any changes

To learn how to run code on R and more about the software and packages, you could see material at: https://education.rstudio.com/learn/beginner/

*Note that indicators in Patient Experiences (PE) and Mental Health PREMs (MP) (that do not require calculations) are not currently eligible for processing with this code.




### (1)  Setup
This section loads the R packages necessary to run the script, sets the working directory of the user in order to be able to access the files. It displays the input data as data.tables as is to verify that the expected files are loaded.
```{r}
### 1. load the packages
# if the packages are not installed yet, you can install them manually by copying the next line (without #) in the console
# install.packages(c("dplyr", "tidyr", "ggplot2", "DT", "shinydashboard", "shiny", "forcats"))

library(dplyr)
library(tidyr)
library(ggplot2)
library(DT)
library(shinydashboard)
library(shiny)
library(forcats)
library(data.table)

### 2. define the path to the folders
# the paths should not be changed and should run if the files are saved correctly
path <- {"../HCQO_validation"}
setwd(path)
file_paths <- list.files("data", pattern = "\\.csv$", full.names = TRUE, recursive = TRUE)

alll <- list()

### 3. preview the input data 
source("functions/load_csv.R")
file_paths <- list.files("data", pattern = "\\.csv$", full.names = TRUE, recursive = TRUE)
lapply(file_paths, load_csv)
```




### (2) Data Validation
This section applies validation checks to the data to verify it. The current validation checks performed are the following:
- Validation Check 1: verifies data format, including column names and values
- Validation Check 2: verifies whether any numerators have no matching denominators
- Validation Check 3: verifies whether there are any rates where the corresponding NUM > DEN
- Validation Check 4: verifies that there are no NA/missing values in otherwise full rows/columns
Validation checks that have passed will be printed below the chunk. If validation check 1 did not pass, it will print out which columns contain incorrect names or values. If validation checks 2-3 do not pass, a table will be printed out with only the rows that made the check fail.

```{r}
source("functions/validate_missing_denom.R")
source("functions/validate_rates.R")
source("functions/validate_format.R")
source("functions/validate_missing_data.R")


all_data <- list() # initialize an empty list to store the data from each file after validation

for (file_path in file_paths) { # iterate over each file path in the list
    df_raw <- load_csv(file_path)
    if (!is.null(df_raw)) { # does not process empty files
        file_name <- basename(file_path)
        print(paste("Processing file:", file_name)) # outputs which file is currently being processed
        
        # Perform validation checks
        col_names_validated <- validate_format(df_raw, file_name) # validation check 1
        validate_missing_denom(df_raw, file_name) # validation check 2
        validate_rates(df_raw, file_name) # validation check 3
        validate_missing_data(df_raw, file_name) # validation check 4
        
        # append data if all validations pass
        if (col_names_validated) {
            all_data[[file_name]] <- df_raw
        }
    }
}

```


### (3) Calculate Rates
This section calls the calculate_indicator function, which infers the indicator, calculates rates and confidence intervals for each indicator type, and merges the result into a single dataframe.
```{r}
source("functions/calculate_indicator.R")

# apply calculate_indicator function to each file path to process data
results_individual <- lapply(file_paths, calculate_indicator)

# combine individual results into a single dataframe for aggregated analysis
results_combined <- dplyr::bind_rows(results_individual)

# remove all white spaces from the MEASURE column to ensure consistency
results_combined$MEASURE <- gsub("\\s", "", results_combined$MEASURE) 

# load the value mapping file with specific data types set for the 'MEASURE' and 'value' columns
value_keys <- read.csv("value_mapping.csv", header = TRUE,
                       colClasses = list('MEASURE' = "character",
                                         'value' = "character"))

# join the processed results with the value mapping based on the 'MEASURE' column to enrich the dataset
merged_df <- merge(results_combined, value_keys, by = "MEASURE", all.x = TRUE)

# add confidence interval values by checking if the existing value is a confidence interval and updating accordingly
merged_df$value.y <- ifelse(merged_df$value.x %in% c('lo_ci', 'up_ci'), yes = merged_df$value.x, no = merged_df$value.y)

# remove the old value column and rename the columns for clarity
merged_df <- merged_df %>% select(-value.x) %>% 
             rename(c(UNIT = value.y, OBS_VALUE = number))

```



## (4) Dashboard
This section displays the results of the calculation in an interactive dashboard. The user can select any specific indicator group/indicator/sex it is interested in via the dropdown lists on the left and can switch between table and graph views at the top of the dashboard.
```{r}
source("functions/dashboard_setup.R") # call the function which contains the dashboard UI settings
shinyApp(ui, server) # display the dashboard
```


## (5) Generate the output file
This section generates a csv file inside HCQO_validation/results which will be named according to the date and time of creation of the file.
```{r}
date.suffix <- format(Sys.time(), "%d%m%Y_%H%M") # extract current date and time

# Create the file name
file.name <- paste0("results_" , date.suffix, ".csv") # name the file with the date and time suffix
file.name.full <- file.path('./results', file.name) # name output path

# Save the file with the date and time suffix
write.csv(merged_df, file = file.name.full, row.names = FALSE)

# Print message once the csv file has been saved
message("Results file successfully saved")

```


